### 2024-03-25 Introduction+SummaryStats.Rmd
Data set is downloaded, cleaned, and initial statistics are done. Considerations for future: there are some missing weeks; how do we make sure to include that there are skips in the data?

```{r}
library(ggplot2)
library(readxl)
library(cowplot)
library(tidyverse)
library(lubridate)
library(dplyr)
```

```{r}
TempData <- read_excel("Temperature2011-2021 (1).xlsx")
TempData <- TempData |>
  mutate(
    `Decimal Years` = date(date_decimal(`Decimal Years`))) |>
  na.omit()

  colnames(TempData) <- c("Date", "Temp", "SD")
```

```{r}
head(TempData)
```

```{r}
ggplot(TempData)+
  geom_point(aes(x = Date, y= Temp)) +
  labs(x = "Year", y = "Temperature") 
```

```{r}
Summary <- TempData |>
  group_by(year(Date)) |>
  summarise(Temp = mean(Temp))
```

```{r}
print(Summary)
```

```{r}
ggplot(Summary)+
  geom_point(aes(x = `year(Date)`, y= Temp)) +
  labs(x = "Year", y = "Average Temperature") 
```

```{r}
SeasonTempData <- TempData |>
  mutate(
    Season = case_when(
      month(Date) == 12 | month(Date) == 1 | month(Date) == 2 ~ "Winter",
      month(Date) == 3 | month(Date) == 4 | month(Date) == 5 ~ "Spring",
      month(Date) == 6 | month(Date) == 7 | month(Date) == 8 ~ "Summer",
      month(Date) == 9 | month(Date) == 10 | month(Date) == 11 ~ "Fall"
    )
  )
```

```{r}
ggplot(SeasonTempData)+
  geom_point(aes(x = Date, y= Temp, color = Season)) +
  labs(x = "Year", y = "Temperature") 
```

```{r}
SummarySeason <- SeasonTempData |>
  group_by(year(Date), Season) |>
  summarise(Temp = mean(Temp))
```

```{r}
print(SummarySeason)
```

```{r}
SummarySeasonSummer <- SummarySeason |>
  filter(Season == "Summer")
print(SummarySeasonSummer)
```

```{r}
SummarySeasonSummer <- SummarySeason |>
  filter(Season == "Winter")
print(SummarySeasonSummer)
```

```{r}
SummarySeasonSummer <- SummarySeason |>
  filter(Season == "Spring")
print(SummarySeasonSummer)
```

```{r}
SummarySeasonSummer <- SummarySeason |>
  filter(Season == "Fall")
print(SummarySeasonSummer)
```