### 2024-04-16 Yilun Upload/YilunCode.Rmd
Yilun new RMD

---
title: "YilunZhuCode"
output: pdf_document
date: "2024-04-15"
always_allow_html: true
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

```{r setup, include=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
Monthly_Data <- read_excel("Monthly Data_Temperature2011-2021.xlsx")
#Transfer Date to only Y/M/D
Monthly_Data$Date <- format(as.Date(Monthly_Data$Date), "%Y/%m")

```

```{r}
ts_Temperature_Monthly <- msts(Monthly_Data$Temperature, 
                           seasonal.periods =c(12),
                           start=c(2011,1)
                           )
#create a subset for training purpose
n_for = 24
ts_Temperature_Monthly_train <- subset(ts_Temperature_Monthly,
                                   end = length(ts_Temperature_Monthly)-n_for)

#create a subset for testing purpose
ts_Temperature_Monthly_test <- subset(ts_Temperature_Monthly,
                                   start = length(ts_Temperature_Monthly)-n_for)
```

```{r}
#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_fit <- nnetar(ts_Temperature_Monthly_train,
                 p=1,
                 P=1,
                 xreg=fourier(ts_Temperature_Monthly_train, K=c(6)))

#NN_for <- forecast(NN_fit, h=6) 
NN_for <- forecast(NN_fit, h=24,xreg=fourier(ts_Temperature_Monthly_train, 
                                          K=c(6),h=24))

#Plot foresting results
autoplot(NN_for) +
  ylab("Temperature C")

#Plot model + observed data
autoplot(ts_Temperature_Monthly_test) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Temperature")

autoplot(ts_Temperature_Monthly) +
  autolayer(NN_for$fitted, series="NN fit",PI=FALSE) +
  autolayer(NN_for$mean, series="NN forecast",PI=FALSE)+
  ylab("Temperature")

checkresiduals(NN_fit)
```

```{r, echo=FALSE}
SSES_seas <- es(ts_Temperature_Monthly_train,model="ZZZ", h=24,holdout=FALSE)
plot(SSES_seas)
checkresiduals(SSES_seas)

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(SSES_seas$fitted, series="SSES Fit",PI=FALSE)+
  autolayer(SSES_seas$forecast, series="SSES Forecast",PI=FALSE)+
  ylab("Temperature") 
```

```{r TBATS, echo=TRUE, message=FALSE, warning=FALSE}
# TBATS can take time to fit
TBATS_fit <- tbats(ts_Temperature_Monthly_train)

TBATS_for <- forecast(TBATS_fit, h=24)

#Plot foresting results
autoplot(TBATS_for) +
  ylab("Temperature")

#Plot test model + observed data
autoplot(ts_Temperature_Monthly_test) +
  autolayer(TBATS_for, series="TBATS Forecast",PI=FALSE)+
  ylab("Temperature")

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(TBATS_fit$fitted.values, series="TBATS Fit",PI=FALSE)+
  autolayer(TBATS_for, series="TBATS Forecast",PI=FALSE)+
  ylab("Temperature")

checkresiduals(TBATS_fit)
```
```{r ETS, echo=TRUE, message=FALSE, warning=FALSE}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(ts_Temperature_Monthly_train,h=24)
#Since stlf did both model+forcastm so h =365 is required to specify the edge of forcast

#Plot foresting results
autoplot(ETS_fit) + ylab("Temperature")

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Temperature")

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(ETS_fit$fitted, series="STL + ETS training",PI=FALSE) +
  autolayer(ETS_fit$mean, series="STL + ETS test",PI=FALSE) +
  ylab("Temperature")

checkresiduals(ETS_fit)
#alternative coding if you don't need to save the objects
#ts_act_power_daily_train %>% stlf(h=365) %>% autoplot() 
#very easy to read/run/implement 
#but you don't have anything stored on your environment

```

```{r ARIMA, echo=TRUE, message=FALSE, warning=FALSE}
#Fit arima model with fourier terms as exogenous regressors
# seasonal = FALSE is the same as P=D=Q=0
# play with K by changing it to K=c(2,2), K=c(2,4), K=c(2,6), etc. The higher the K the longer it will take to converge, because R will try more models.

ARIMA_Four_fit <- auto.arima(ts_Temperature_Monthly_train, 
                             seasonal=TRUE, 
                             lambda=0,
                             xreg=fourier(ts_Temperature_Monthly_train, 
                                          K=c(6))
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(ts_Temperature_Monthly_train,
                                        K=c(6),
                                        h=24),
                           h=24
                           )

#Plot foresting results
autoplot(ARIMA_Four_for) + ylab("Temperature")

#Plot model + observed data
autoplot(ts_Temperature_Monthly) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Temperature")

```
```{r}

#Model 1: STL + ETS
ETS_scores <- accuracy(ETS_fit$mean,ts_Temperature_Monthly_test)

#Model 2: ARIMA + Fourier 
ARIMA_scores <- accuracy(ARIMA_Four_for$mean,ts_Temperature_Monthly_test)

# Model 3:  TBATS 
TBATS_scores <- accuracy(TBATS_for$mean,ts_Temperature_Monthly_test)

# Model 4:  Neural Network 
NN_scores <- accuracy(NN_for$mean,ts_Temperature_Monthly_test)

# Model 5:  SSES 
SSES_scores <- accuracy(SSES_seas$forecast,ts_Temperature_Monthly_test)

```

```{r}
#create data frame
scores <- as.data.frame(
  rbind(ETS_scores, ARIMA_scores, TBATS_scores, NN_scores)
  )
row.names(scores) <- c("STL+ETS", "ARIMA+Fourier","TBATS","NN","SSES")

#choose model with lowest RMSE
best_model_index <- which.min(scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(scores[best_model_index,]))                       
```

```{r}
autoplot(ts_Temperature_Monthly) +
  autolayer(ETS_fit, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_Four_for, PI=FALSE, series="ARIMA + Fourier") +
  autolayer(TBATS_for,PI=FALSE, series="TBATS") +
  autolayer(NN_for,PI=FALSE, series="NN") +
  xlab("Monthly") + ylab("Water Temperature") +
  guides(colour=guide_legend(title="Forecast"))

```